Modelling last glacial cycle ice dynamics in the Alps

### 2017-07-13 articles/cycle/alps-cycle.tex
Move palaeo forcing methods to dedicated section.

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Climate forcing}
\label{sec:atm}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    \idea{move paleo-climate forcing info to the next section.}

    Climate forcing driving ice sheet simulations consists of a~present-day
    monthly climatology, $\{T_{\mathrm{m}0}, P_{\mathrm{m}0}\}$, where
    temperatures are modified by offset time series, ${\Delta}T_{\text{TS}}$,
    and lapse-rate corrections, ${\Delta}T_{\text{LR}}$:
    %
    \begin{align}
      T_{\mathrm{m}}(t, x, y) = T_{\mathrm{m}0}(x, y) +
                                {\Delta}T_{\text{TS}}(t) +
                                {\Delta}T_{\text{LR}}(t, x, y) \,, \\
    \end{align}
    %
    The present-day monthly climatology was bilinearly interpolated from
    near-surface air temperature and precipitation rate fields from
    \citep[WorldClim;][]{Hijmans.etal.2005}, representative of the period 1960
    to 1990. Modern climate of the European Alps is characterised by a
    latitudinal gradient in summer air temperatures, and a longitudinal
    gradient in winter precipitation (Fig.~\ref{fig:inputs}). WorldClim data
    were selected as an input to the ice sheet model because they incorporate
    observations from the dense weather station network of central Europe.
    Besides, last glacial cycle alpine glaciers did not extend over marine
    areas where WorldClim data is missing. Finally, WorldClim data were
    previously used as climate forcing for PISM to model the LGM extent of the
    former Cordilleran ice sheet in good agreement with geological evidence
    along the southern margin \citep{Seguinot.etal.2014} were weather station
    density is lower than in the Alps.

@@

    Table~\ref{tab:records} -- Palaeo-climate records.\\
    Fig.~\ref{fig:timeseries} -- Low-resolution time series.\\
    Fig.~\ref{fig:footprints} -- Low-resolution ice cover.

    \begin{itemize}
    \item We conducted a 5\,km-resolution sensitivity study.
    \item We used three palaeo-temperature records (Table~\ref{tab:records}).
    \item All records scaled to fit the LGM extent (Table~\ref{tab:records}).
    \end{itemize}

    \begin{itemize}
    \item Some simulations use reduced precipitation (Table~\ref{tab:records}).
    \end{itemize}

>>>>

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Reference climate forcing}
\label{sec:atm}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    Climate forcing driving ice sheet simulations consists of a~present-day
    monthly climatology, $\{T_{\mathrm{m}0}, P_{\mathrm{m}0}\}$, where
    temperatures are modified by offset time series, ${\Delta}T_{\text{TS}}$,
    and lapse-rate corrections, ${\Delta}T_{\text{LR}}$:
    %
    \begin{align}
      T_{\mathrm{m}}(t, x, y) = T_{\mathrm{m}0}(x, y) +
                                {\Delta}T_{\text{TS}}(t) +
                                {\Delta}T_{\text{LR}}(t, x, y) \,,
    \end{align}
    %
    The present-day monthly climatology was bilinearly interpolated from
    near-surface air temperature and precipitation rate fields from
    \citep[WorldClim;][]{Hijmans.etal.2005}, representative of the period 1960
    to 1990. Modern climate of the European Alps is characterised by a
    latitudinal gradient in summer air temperatures, and a longitudinal
    gradient in winter precipitation (Fig.~\ref{fig:inputs}). WorldClim data
    were selected as an input to the ice sheet model because they incorporate
    observations from the dense weather station network of central Europe.
    Besides, last glacial cycle alpine glaciers did not extend over marine
    areas where WorldClim data is missing. Finally, WorldClim data were
    previously used as climate forcing for PISM to model the LGM extent of the
    former Cordilleran ice sheet in good agreement with geological evidence
    along the southern margin \citep{Seguinot.etal.2014} were weather station
    density is lower than in the Alps.

@@
